---
title: "Figure S13: Differential expression analysis between tissues"
author: "Karol Nowicki-Osuch, based on Nils Eling's code"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script focuses on detecting the genes that are differentially expressed between BE and BSCJ, NG and NSCJ, NE and NSCJ and NE and BSCJ

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(DropletUtils)
library(openxlsx)
library(Rtsne)
library(pheatmap)
library(RColorBrewer)
library(edgeR)
library(viridis)
source("../../Analysis/Functions/auxiliary.R")

sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

sce <- sce[,sce$include]
```

# Visualize NE, NG, NSCJ, BE and BSCJ samples

Here we perform differential expression (DE) analysis between the epithelial tissues of the NG, NSCJ, BE, BSCJ and NE samples.

```{r}
# Select tissues to test
cur_sce <- sce[,sce$include & sce$Tissue %in% c("NE", "NG", "BE", "NSCJ", "BSCJ")]

# Perform batch correction
sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")

# Order sce objects for batch correction
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

n <- names(sce.list)
sce.list <- sce.list[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
                       which(grepl("NE_out", n)), which(grepl("NG_out", n)),
                       which(grepl("BE_out", n)))]


corrected <- batch.correction(sce.list)
cur_sce <- do.call("cbind", sce.list)

# Compute new tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)
reducedDims(cur_sce)$TSNE <- tsne$Y

# Visualize tSNEs
# Tissue

p.all.cells <- ggplot(data.frame(tsne1 = reducedDims(cur_sce)$TSNE[,1],
                  tsne2 = reducedDims(cur_sce)$TSNE[,2],
                  tissue = colData(cur_sce)$Tissue)) + 
  geom_point(aes(tsne1, tsne2), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue), size = 0.5) + 
  scale_color_manual(values = metadata(cur_sce)$colour_vector) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))
p.all.cells
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S13/All_cells.pdf", p.all.cells, width = 7, height = 5, useDingbats = FALSE)

# Tissue typse
p.all.cells.tissue_type <- ggplot(data.frame(tsne1 = reducedDims(cur_sce)$TSNE[,1],
                  tsne2 = reducedDims(cur_sce)$TSNE[,2],
                  tissue = colData(cur_sce)$tissue_type)) +
  geom_point(aes(tsne1, tsne2), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue), size = 0.5) + 
  scale_color_manual(values = c("darkblue", "grey40", "darkred" )) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))
p.all.cells.tissue_type
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S13/Tissue_type.pdf", p.all.cells.tissue_type, width = 7, height = 5, useDingbats = FALSE)

# Cell type
colour_vector <- vector(length = length(unique(cur_sce$cell_type)))
names(colour_vector) <- unique(cur_sce$cell_type)
colour_vector["Superficial"] <- colorRampPalette(c("white", "dark red"))(10)[10]
colour_vector["Basal"] <- colorRampPalette(c("white", "dark red"))(10)[4]
colour_vector["Suprabasal"] <- colorRampPalette(c("white", "dark red"))(10)[6]
colour_vector["Intermediate"] <- colorRampPalette(c("white", "dark red"))(10)[8]

colour_vector["Foveolar_differentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[10]
colour_vector["Endocrine"] <- colorRampPalette(c("white", "dark blue"))(10)[10]
colour_vector["Undifferentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[3]
colour_vector["Foveolar_Intermediate"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[6]
colour_vector["Chief"] <- colorRampPalette(c("white", "dark blue"))(10)[6]
colour_vector["Parietal"] <- colorRampPalette(c("white", "dark blue"))(10)[3]

colour_vector["Goblet"] <- colorRampPalette(c("white", "seagreen4"))(10)[10]
colour_vector["Columnar_differentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[10]
colour_vector["Columnar_Undifferentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[3]
colour_vector["Columnar_Intermediate"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[6]

colour_vector["Immune"] <- "grey40"
colour_vector["Stromal"] <- "grey60"

colour_vector["MUC5B_cells"] <- "saddlebrown"
colour_vector["KRT7_cells"] <- "burlywood3"
colour_vector["KRT5.KRT7_cells"] <- "burlywood4"
colour_vector["KRT5_cells"] <- "orange"

p.all.cells.cell_type <- ggplot(data.frame(tsne1 = reducedDims(cur_sce)$TSNE[,1],
                  tsne2 = reducedDims(cur_sce)$TSNE[,2],
                  tissue = colData(cur_sce)$cell_type)) +
  geom_point(aes(tsne1, tsne2), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue), size = 0.5) + 
  scale_color_manual(values = colour_vector) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))

p.all.cells.cell_type

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S13/Cell_type.pdf", p.all.cells.cell_type, width = 7, height = 5, useDingbats = FALSE)

# Visualize patient
# Colouring for patient
colour_vector <- vector(length = length(unique(cur_sce$Patient)))
names(colour_vector) <- unique(cur_sce$Patient)
colour_vector <- brewer.pal(12, "Set3")

p.all.cells.patient <- ggplot(data.frame(tsne1 = reducedDims(cur_sce)$TSNE[,1],
                  tsne2 = reducedDims(cur_sce)$TSNE[,2],
                  tissue = colData(cur_sce)$Patient)) +
  geom_point(aes(tsne1, tsne2), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue), size = 0.5) + 
  scale_color_manual(values = colour_vector) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))
p.all.cells.patient

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S13/Patient.pdf", p.all.cells.patient, width = 7, height = 5, useDingbats = FALSE)
```

# Differential expression analysis

Here we berform differential expression analysis between NE and squamous cells of GOJ, GC and columnar cells of GC, NE and squamous cells of SCJ, BE and columnar cells of SCJ.
For this, we will sum the raw counts for each gene across all cells per tissue and per patient. 
We then use edgeR on these pseudo-bulk samples to test for changes in mean expression.

```{r}
# Define conditions to test
# First two slots are tissues and last slot is cell-types to compare
cond <- list(c("NE", "NSCJ", "Basal-Intermediate-Superficial-Suprabasal", "Squamous"),
             c("NG", "NSCJ", "Foveolar_Intermediate-Foveolar_differentiated-Endocrine-Undifferentiated-Chief", "Columnar"),
             c("BE", "BSCJ", "Columnar_differentiated-Columnar_Intermediate-Columnar_Undifferentiated-Endocrine", "Columnar"),
             c("NE", "BSCJ", "Basal-Intermediate-Superficial-Suprabasal", "Squamous"))

for(i in 1:length(cond)){
  tissue1 <- cond[[i]][1]
  tissue2 <- cond[[i]][2]
  cell_types <- unlist(strsplit(cond[[i]][3], "-"))
  tissue_type <- cond[[i]][4]
  # Select cells
  sce_to_test <- cur_sce[,(cur_sce$Tissue == tissue1 | 
                             cur_sce$Tissue == tissue2) &
                           cur_sce$cell_type %in% cell_types &
                           cur_sce$tissue_type == tissue_type]
  
  # Remove non-expressed genes 
  sce_to_test <- sce_to_test[Matrix::rowSums(counts(sce_to_test)) > 0,]
  
  # Peform DE
  cur_out <- DE.edgeR(sce = sce_to_test,
                      conditions = sce_to_test$Tissue,
                      covariate = sce_to_test$Patient,
                      lfc = 0.5,
                      FDR = 0.1)
  
  # Calculate the percentage of DE genes
  print(nrow(do.call("rbind", cur_out))/nrow(sce_to_test))
}

# NE vs NSCJ: <0.01%
# GC vs NSCJ: 0.51%
# BE vs BSCJ: 0.00%
# NE vs BSCJ: 0.37%
```

```{r}
sessionInfo()
```