---
title: "IPA_output BE diff cells vs BE undiff"
author: "Karol"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

This is output of the standard analysis of IPA. I am printing p-values for the top 20 upstream regulators that are transcriptional regulators. In the IPA, I did standard analysis with the only setting changed to Direct Interactions.

IPA:
Content version: 49932394 (Release Date: 2019-11-14)

```{r setup, include=FALSE}
# export RSTUDIO_PANDOC=/usr/lib/rstudio-server/bin/pando
# Rscript -e "library(rmarkdown); render('FILE_NAME.Rmd')"
knitr::opts_chunk$set(	echo=TRUE,
                       results='asis',
                       fig.align='center')
```

```{r setup, , echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
library(ggplot2)
files.dir<-"/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures//S16/"
location<-paste0(files.dir, "/Data/IPA")

```

## Figure 16
Here I am comparing the differentiated cells of stomach with differentiated cells of BE using IPA.
Content version: 51963813 (Release Date: 2020-03-11)



```{r process_16, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 14, fig.width= 10}
#Read data
all.data<-read.delim(file=paste0(location,"/IPA_S7_Diff_BEvNG_upstream.txt"),  skip = 2)
#Calc log of p-values
all.data$logp<- -log10(all.data$p.value.of.overlap)
# all.data$p.adj<- p.adjust(all.data$p.value.of.overlap, method = "BH")
# all.data$logq<- -log10(all.data$p.adj)
all.data$logq<- -log10(all.data$B.H.corrected.p.value)


all.data<-all.data[rev(order(all.data$logq)),]
all.data$no.genes<-sapply(as.character(all.data$Target.Molecules.in.Dataset), function(x) {length(unlist(strsplit(x, split = ",")))})

type = "transcription regulator" 
max=20
data.tmp<-all.data[all.data$Molecule.Type==type,]
data.tmp$Activation.z.score[is.na(data.tmp$Activation.z.score)]<-0
#keep only data with the positive z-score
data.tmp<-data.tmp[data.tmp$Activation.z.score>0,]
#check if there are 20 elements in the group
if(nrow(data.tmp)<=20){max<-nrow(data.tmp)}
data.tmp<-data.tmp[data.tmp$Molecule.Type==type,][1:max,]

p<-ggplot(data=data.tmp, aes(x=Upstream.Regulator, y=logq,  label = no.genes))+
  geom_bar(stat = "identity" , color = "black", fill = "black", width = 0.5)+coord_flip()           
p<-p  + scale_x_discrete(limits=rev(data.tmp$Upstream.Regulator))
p<-p+ggtitle(label = type)+ylab(label = "-Log10(q-value)") + xlab(label = type)+geom_text(hjust = -.5)+ theme_bw()

p

ggsave(plot = p, file = paste0(files.dir, "/Diff_BEvNG_Upregulated.pdf"),  width = 8, height = 10, useDingbats=FALSE)


type = "transcription regulator" 
max=20
data.tmp<-all.data[all.data$Molecule.Type==type,]

data.tmp$Activation.z.score[is.na(data.tmp$Activation.z.score)]<-0
#keep only data with the positive z-score
data.tmp<-data.tmp[data.tmp$Activation.z.score>0,]

#keep only data with with qval <0.01
data.tmp<-data.tmp[data.tmp$B.H.corrected.p.value<0.01,]

data.tmp <- data.tmp[rev(order(data.tmp$Activation.z.score)),]

#check if there are 20 elements in the group
if(nrow(data.tmp)<=max){max<-nrow(data.tmp)}
data.tmp<-data.tmp[data.tmp$Molecule.Type==type,][1:max,]
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
data.tmp$logq <- specify_decimal(data.tmp$logq, 2)

p<-ggplot(data=data.tmp, aes(x=Upstream.Regulator, y=Activation.z.score,  label = logq))+
  geom_bar(stat = "identity" , color = "black", fill = "black", width = 0.5)+coord_flip()           
p<-p  + scale_x_discrete(limits=rev(data.tmp$Upstream.Regulator))
p<-p+ggtitle(label = type)+ylab(label = "Activation Z-score") + xlab(label = type)+geom_text(hjust = -.1)+ theme_bw()

p


```


## End Matters

To finish get session info:
```{r Endnote, echo=FALSE}
sessionInfo()
```
