---
title: "Figure S8: Computational processing of all data"
author: "Karol Nowicki-Osuch based on Nils Eling's code"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script visualizes the filtered data obtained after running the Quality_control, Normalization and Further_processing scripts. 
We will first plot these cells to highlight the patient effect and afterwards perform batch-correction.
Finally, we will mark the high-quality batches and cells.

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(DropletUtils)
library(openxlsx)
library(Rtsne)
#library(ape)
library(viridis)
library(umap)
source("../../Analysis/Functions/auxiliary.R")
```

# Dimensionality reduction

We will first visualize the uncorrected data. I will also create a circular dendrogram on the same selected 1000 genes

```{r}
# Read in all datafiles
sce.list <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_sce_filtered.rds")

# Remove all unecessary rowData prior to merging
sce.list <- lapply(sce.list, function(n){
  rowData(n) <- rowData(n)[,1:3]
  n
})

#merge datasets
sce.all <- do.call("cbind", sce.list)
# # merged data already exists in All_corrected_sce_filtered.rds. It also has clustering results.
# sce.all <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Compute the top 1000 shared highly variable genes - use the new code for scran
HVG.genes <- lapply(sce.list, function(n){
  HVG <- modelGeneVar(n)
  HVG
})
HVG.df <- do.call("combineVar", HVG.genes)
HVG.df <- HVG.df[order(HVG.df$bio, decreasing = TRUE),]
HVgenes <- rownames(HVG.df)[1:1000]



# # Compute the top 1000 shared highly variable genes
# HVG.genes <- lapply(sce.list, function(n){
#   HVG <- trendVar(n, use.spikes = FALSE)
#   decomposeVar(n, HVG)
# })
# HVG.df <- do.call("combineVar", HVG.genes)
# HVG.df <- HVG.df[order(HVG.df$bio, decreasing = TRUE),]
# HVgenes <- rownames(HVG.df)[1:1000]

# Compute tsne
set.seed(123456)
tsne <- Rtsne(t(as.matrix(logcounts(sce.all)[HVgenes,])), perplexity = 250, pca = FALSE)
UMAP <-umap(t(as.matrix(logcounts(sce.all)[HVgenes,])))
# Create annotation and colour vector
annot <- paste(colData(sce.all)$Tissue, colData(sce.all)$Patient, sep = "_")
annot.col <- vector(length = length(unique(annot)))
names(annot.col) <- unique(annot)
annot.col[grepl("NE", names(annot.col))] <- colorRampPalette(c("white", "darkred"))(21)[c(5,7,9,11,13,15,17,19,21)]
annot.col[grepl("NSCJ", names(annot.col))] <- colorRampPalette(c("white", "#F39B7FFF"))(21)[c(7,9,11,13,15,17,19,21)]
annot.col[grepl("BSCJ", names(annot.col))] <- colorRampPalette(c("white", "#DC0000FF"))(17)[c(13,15,17)]
annot.col[grepl("BE", names(annot.col))] <- colorRampPalette(c("white", "#00A087FF"))(17)[c(6,9,13,16)]
annot.col[grepl("ND", names(annot.col))] <- colorRampPalette(c("white", "#3C5488FF"))(17)[c(5,9,13,17)]
annot.col[grepl("NG", names(annot.col))] <- colorRampPalette(c("white", "#4DBBD5FF"))(25)[c(5,7,9,11,13,15,17,19,21,23,25)]
annot.col[grepl("SMG", names(annot.col))] <- colorRampPalette(c("white", "#B09C85FF"))(17)[c(5,9,13)]
annot.col[grepl("GM", names(annot.col))] <- colorRampPalette(c("white", "#00A087FF"))(17)[c(5)]

# Plot tsne
tsne.uncorrected <- ggplot(data.frame(tsne1 = tsne$Y[,1], 
                                      tsne2 = tsne$Y[,2], 
                                      Sample = annot)) + 
  geom_point(aes(tsne1, tsne2, colour = Sample), size = 0.5) + 
  scale_colour_manual(values = annot.col) + 
  theme_minimal()
tsne.uncorrected

# Plot umap
umap.uncorrected <- ggplot(data.frame(umap1 = UMAP$layout[,1], 
                                      umap2 = UMAP$layout[,2], 
                                      Sample = annot)) + 
  geom_point(aes(umap1, umap2, colour = Sample), size = 0.5) + 
  scale_colour_manual(values = annot.col) + 
  theme_minimal()
umap.uncorrected

#show expression of immune and stromal cells

#Goblet cells
gene1<-"MUC2"
GOBcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                            tSNE2 = tsne$Y[,2],
                            gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
GOBcells

#Goblet cells
gene1<-"CHGA"
CHGAcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                             tSNE2 = tsne$Y[,2],
                             gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
CHGAcells

#Stromal cells
gene1<-"VIM"
VIMcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                            tSNE2 = tsne$Y[,2],
                            gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
VIMcells

# Immune cells
gene1<-"PTPRC"
PTPRCcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
PTPRCcells


ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_uncorrected.pdf", tsne.uncorrected, width = 12, height = 7, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_uncorrected_VIMexpression.pdf", VIMcells, width = 9, height = 7, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_uncorrected_PTPRCexpression.pdf", PTPRCcells, width = 9, height = 7, useDingbats = FALSE)






```

# Batch correction across tissues

We will now perform batch correction across all samples. 
For this, we will use the samples that contain a mix of different tissues first.

```{r}
# Order sce objects for batch correction



sce.list <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_sce_filtered.rds")

# Order sce objects for batch correction
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

n <- names(sce.list)
sce.list <- sce.list[c(which(grepl("NSCJ", n)), which(grepl("BSCJ", n)), 
                       which(grepl("NE", n)), which(grepl("NG", n)),
                       which(grepl("BE", n)), which(grepl("ND", n)),
                       which(grepl("SMG", n)), which(grepl("GM", n)))]

# Remove all unecessary rowData prior to merging
sce.list <- lapply(sce.list, function(n){
  rowData(n) <- rowData(n)[,1:3]
  n
})

#merge datasets
sce.all <- do.call("cbind", sce.list)


# Batch correction
corrected <- batch.correction(sce.list)


# Compute new tSNE
# set.seed(11111)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 250)
UMAP <- umap(t(corrected))
# Visualize batches
annot <- paste(colData(sce.all)$Tissue, colData(sce.all)$Patient, sep = "_")

tsne.corrected <- ggplot(data.frame(tsne1 = tsne$Y[,1], tsne2 = tsne$Y[,2], 
                                    Sample = annot)) + 
  geom_point(aes(tsne1, tsne2, colour = Sample)) + scale_colour_manual(values = annot.col) + theme_minimal()
tsne.corrected 


umap.corrected <- ggplot(data.frame(umap1 = UMAP$layout[,1], umap2 = UMAP$layout[,2], 
                                    Sample = annot)) + 
  geom_point(aes(umap1, umap2, colour = Sample)) + scale_colour_manual(values = annot.col) + theme_minimal()
umap.corrected 

#show expression of immune and stromal cells

#Goblet cells
gene1<-"MUC2"
GOBcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                            tSNE2 = tsne$Y[,2],
                            gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
GOBcells


#Stromal cells
gene1<-"VIM"
VIMcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                            tSNE2 = tsne$Y[,2],
                            gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
VIMcells

# Immune cells
gene1<-"PTPRC"
PTPRCcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
PTPRCcells
# CHGA cells
gene1<-"CHGA"
CHGAcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                             tSNE2 = tsne$Y[,2],
                             gene = logcounts(sce.all)[rowData(sce.all)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 0.5) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
CHGAcells


ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_corrected.pdf", tsne.corrected, width = 12, height = 7, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_corrected_VIMexpression.pdf", VIMcells, width = 9, height = 7, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_corrected_PTPRCexpression.pdf", PTPRCcells, width = 9, height = 7, useDingbats = FALSE)




```

# High quality samples

In the next step, we will highlight the high quality samples. 
The exclusion criteria is explained in Suppementary Note 1.

```{r}
# Read in the data of high quality samples
sce.high.quality <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")
# annot.high.quality <- sapply(names(sce.high.quality), function(n){paste(unlist(strsplit(n, "_"))[c(3,1)], collapse = "_")})
annot.high.quality <- unique(paste0(colData(sce.high.quality)$Tissue,"_",colData(sce.high.quality)$Patient))
# Grey out the cells that we are not using
annot.col.high.quality <- annot.col
annot.col.high.quality[!(names(annot.col.high.quality) %in% annot.high.quality)] <- "grey20"

# Dataframe for visualization
df <- data.frame(tsne1 = tsne$Y[,1], tsne2 = tsne$Y[,2], 
                 Sample = annot,
                 include = paste(sce.all$Tissue, sce.all$Patient, sep = "_") %in%
                   annot.high.quality)

tsne.corrected.high.quality <- ggplot() + 
  geom_point(data = df[df$include,],
             aes(tsne1, tsne2, colour = Sample)) +
  geom_point(data = df[!df$include,],
             aes(tsne1, tsne2, colour = Sample), shape = 1, alpha = 0.5) +
  scale_colour_manual(values = annot.col.high.quality) + 
  theme_minimal()
tsne.corrected.high.quality
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_corrected_highQuality.pdf", tsne.corrected.high.quality, width = 12, height = 7, useDingbats = FALSE)
```

# View clusters that were removed due to possible contaminations

In the final step, we remove groups of cells that present contaminations when sampling the tissues.
For example these are squamous oesophagus cells found in the gastric sample.

```{r}
sce.final <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Create new colour vector
annot <- paste(sce.all$Tissue, sce.all$Patient, sep = "_")
annot[!(annot %in% annot.high.quality)] <- "exclude"
annot[match(paste(sce.final$Sample[!sce.final$include], sce.final$Barcode[!sce.final$include], sep = "_"), paste(sce.all$Sample, sce.all$Barcode, sep = "_"))] <- "exclude"

# Grey out the cells that we are not using
annot.col.high.quality <- c(annot.col, "grey20")
names(annot.col.high.quality)[length(annot.col.high.quality)] <- "exclude"

# Dataframe for visualization
df <- data.frame(tsne1 = tsne$Y[,1], tsne2 = tsne$Y[,2], 
                 Sample = annot)

tsne.corrected.final <- ggplot() + 
  geom_point(data = df[df$Sample != "exclude",],
             aes(tsne1, tsne2, colour = Sample)) +
  geom_point(data = df[df$Sample == "exclude",],
             aes(tsne1, tsne2, colour = Sample), alpha = 0.25) +
  
  scale_colour_manual(values = annot.col.high.quality) + 
  theme_minimal()
tsne.corrected.final
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S8/All_data_corrected_final.pdf", tsne.corrected.final, width = 12, height = 7, useDingbats = FALSE)
```


# End Matter

To finish get session info:

```{r Endnote, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
sessionInfo()
```
