---
title: "Figure S10: Entropy measures in all tissues"
author: "Karol Nowicki-Osuch, based on Nils Eling's code"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, we compute the entropy per cell for duodenum and normal oesophagus cells as a control for this measure to detect stem-cell-like phenotypes.

# Read in the data

```{r libraries, message=FALSE}
library(Rtsne)
library(viridis)
source("../../Analysis/Functions/auxiliary.R")

sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

sce <- sce[,sce$include]
```

# Entropy in ND samples

Here, we compute the entropy in the duodenum samples.

```{r}
# Select ND cells
cur_sce <- sce[,sce$Tissue == "ND"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <- vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_B-cells"] <- "grey30"
colour_vector["Immune_Macrophages"] <- "grey40"
colour_vector["Stromal_CALD1_cells"] <- "grey60"
colour_vector["Stromal_GNG11_cells"] <- "grey70"

colour_vector["Enterocytes_Intermediate"] <- colorRampPalette(c("white", "#3C5488FF"))(10)[6]
colour_vector["Enterocytes_differentiated"] <- colorRampPalette(c("white", "#3C5488FF"))(10)[10]
colour_vector["Undifferentiated_Dividing"] <- colorRampPalette(c("white", "#3C5488FF"))(10)[3]
colour_vector["Undifferentiated"] <- colorRampPalette(c("white", "#3C5488FF"))(10)[1]
colour_vector["Goblet"] <- colorRampPalette(c("white", "seagreen4"))(10)[10]
colour_vector["Paneth"] <- colorRampPalette(c("white", "seagreen4"))(10)[6]
colour_vector["Endocrine_CHGA"] <- colorRampPalette(c("white", "dark blue"))(10)[10]


tsne.D2 <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.D2
# Colour by entropy
tsne.D2.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.D2.entropy
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures//Supplementary_figures/S10/D2_tsne.pdf", tsne.D2, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures//Supplementary_figures/S10/D2_tsne_entropy.pdf", tsne.D2.entropy, width = 6, height = 5, useDingbats = FALSE)
```

# Entropy in NE samples

Here, we compute the entropy in the normal oesophagus samples.

```{r}
# Select NE cells
cur_sce <- sce[,sce$Tissue == "NE"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <- vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_Macrophages"] <- "grey40"

colour_vector["Superficial"] <- colorRampPalette(c("white", "dark red"))(10)[10]
colour_vector["Basal"] <- colorRampPalette(c("white", "dark red"))(10)[4]
colour_vector["Suprabasal"] <- colorRampPalette(c("white", "dark red"))(10)[6]
colour_vector["Intermediate"] <- colorRampPalette(c("white", "dark red"))(10)[8]
colour_vector["Suprabasal_Dividing"] <- colorRampPalette(c("white", "dark red"))(10)[1]

tsne.NE <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.NE
# Colour by entropy
tsne.NE.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.NE.entropy
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NE_tsne.pdf", tsne.NE, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NE_tsne_entropy.pdf", tsne.NE.entropy, width = 6, height = 5, useDingbats = FALSE)
```

# Entropy in BE samples

```{r}
# Select BE cells
cur_sce <- sce[,sce$Tissue == "BE"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <- vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_B-cells"] <- "grey30"
colour_vector["Immune_Macrophages"] <- "grey40"
colour_vector["Stromal_CALD1_cells"] <- "grey60"
colour_vector["Stromal_GNG11_cells"] <- "grey70"

colour_vector["Columnar_Undifferentiated"] <- "white"
colour_vector["Columnar_Undifferentiated_Dividing"] <- colorRampPalette(c("white", "#00A087FF"))(10)[2]
colour_vector["Columnar_Intermediate"] <- colorRampPalette(c("white", "#00A087FF"))(10)[4]
colour_vector["Columnar_differentiated"] <- colorRampPalette(c("white", "#00A087FF"))(10)[6]
colour_vector["Endocrine_NEUROG3"] <- colorRampPalette(c("white", "dark green"))(10)[5]
colour_vector["Goblet"] <- colorRampPalette(c("white", "dark green"))(10)[10]

tsne.BE <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.BE
# Colour by entropy
tsne.BE.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.BE.entropy
# Visualize OLFM4
tsne.BE.OLFM4 <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  OLFM4 = log2(normcounts(cur_sce)[rowData(cur_sce)$Symbol == "OLFM4",] + 1))) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = OLFM4)) +
  scale_color_viridis() + theme_minimal()
tsne.BE.OLFM4
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures//Supplementary_figures/S10/BE_tsne.pdf", tsne.BE, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures//Supplementary_figures/S10/BE_tsne_entropy.pdf", tsne.BE.entropy, width = 6, height = 5, useDingbats = FALSE)
```

# Entropy in NG samples

```{r}
# Select NG cells
cur_sce <- sce[,sce$Tissue == "NG"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <- vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_B-cells"] <- "grey30"
colour_vector["Immune_Macrophages"] <- "grey40"
colour_vector["Stromal_CALD1_cells"] <- "grey60"
colour_vector["Stromal_GNG11_cells"] <- "grey70"


colour_vector["Foveolar_differentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[10]
colour_vector["Undifferentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[3]
colour_vector["Foveolar_Intermediate"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[6]
colour_vector["Chief"] <- colorRampPalette(c("white", "dark blue"))(10)[6]
colour_vector["Parietal"] <- colorRampPalette(c("white", "dark blue"))(10)[3]
colour_vector["Undifferentiated_Dividing"] <- colorRampPalette(c("white", "dark blue"))(10)[1]

colour_vector["Endocrine_GHRL"] <- colorRampPalette(c("white", "#5900D8FF"))(10)[10]
colour_vector["Endocrine_CHGA"] <- colorRampPalette(c("white", "#5900D8FF"))(10)[7]
colour_vector["Endocrine_NEUROD1"] <- colorRampPalette(c("white", "#5900D8FF"))(10)[4]


tsne.GC <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.GC
# Colour by entropy
tsne.GC.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.GC.entropy
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NG_tsne.pdf", tsne.GC, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NG_tsne_entropy.pdf", tsne.GC.entropy, width = 6, height = 5, useDingbats = FALSE)
```

# Entropy in SMG samples

```{r}
# Select SMG cells
cur_sce <- sce[,sce$Tissue == "SMG"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <-  vector(length = length(unique(cur_sce$cell_type)))
names(colour_vector) <- unique(cur_sce$cell_type)
colour_vector["Immune"] <- "grey40"
colour_vector["Unknown.Doublets"] <- "white"
colour_vector["Mucous"] <- "saddlebrown"
colour_vector["Oncocytes"] <- "burlywood3"
colour_vector["Stromal"] <- "grey60"
colour_vector["Duct_Intercalating"] <- "burlywood4"
colour_vector["Squamous_Esophagus"] <- "black"
colour_vector["Myo-epithelial"] <- "brown"

tsne.SMG <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.SMG
# Colour by entropy
tsne.SMG.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.SMG.entropy
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/SMG_tsne.pdf", tsne.SMG, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/SMG_tsne_entropy.pdf", tsne.SMG.entropy, width = 6, height = 5, useDingbats = FALSE)
```

# Entropy in NSCJ samples

```{r}
# Select NSCJ cells
cur_sce <- sce[,sce$Tissue == "NSCJ"]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = FALSE)

cur_counts <- as.matrix(normcounts(cur_sce))

# Calculate entropy
probs   <- t(t(cur_counts)/colSums(cur_counts))
logprobs <- log(probs)
logprobs[logprobs == -Inf] <- 0
logprobs.1 <- probs*logprobs
entropy <- -colSums(probs*logprobs)

# Entropy is normalized based on the number of genes
entropy <- entropy/log(nrow(cur_counts))

sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)

# Create colour vector
colour_vector <-  vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_B-cells"] <- "grey30"
colour_vector["Immune_Macrophages"] <- "grey40"
colour_vector["Stromal_CALD1_cells"] <- "grey60"
colour_vector["Stromal_GNG11_cells"] <- "grey70"

colour_vector["Foveolar_differentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[10]
colour_vector["Undifferentiated"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[3]
colour_vector["Foveolar_Intermediate"] <- colorRampPalette(c("white", "#4DBBD5FF"))(10)[6]

colour_vector["Endocrine_GHRL"] <- colorRampPalette(c("white", "#5900D8FF"))(10)[10]
colour_vector["Endocrine_CHGA"] <- colorRampPalette(c("white", "#5900D8FF"))(10)[7]

colour_vector["Superficial"] <- colorRampPalette(c("white", "dark red"))(10)[10]
colour_vector["Basal"] <- colorRampPalette(c("white", "dark red"))(10)[4]
colour_vector["Suprabasal"] <- colorRampPalette(c("white", "dark red"))(10)[6]
colour_vector["Intermediate"] <- colorRampPalette(c("white", "dark red"))(10)[8]
colour_vector["Suprabasal_Dividing"] <- colorRampPalette(c("white", "dark red"))(10)[1]

colour_vector["MUC5B_cells"] <- "saddlebrown"
colour_vector["KRT7_cells"] <- "burlywood3"
colour_vector["KRT5.KRT7_cells"] <- "burlywood4"
colour_vector["KRT5_cells"] <- "orange"

tsne.NSCJ <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  scale_color_manual(values = colour_vector) + theme_minimal()
tsne.NSCJ
# Colour by entropy
tsne.NSCJ.entropy <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  entropy = entropy)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) + 
  geom_point(aes(tSNE1, tSNE2, colour = entropy)) +
  scale_color_gradientn(colours = magma(100)) + theme_minimal()
tsne.NSCJ.entropy
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NSCJ_tsne.pdf", tsne.NSCJ, width = 6, height = 5, useDingbats = FALSE)
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S10/NSCJ_tsne_entropy.pdf", tsne.NSCJ.entropy, width = 6, height = 5, useDingbats = FALSE)
```
