---
title: "Figure S12: Similarity between BE and all other tissues"
author: "Karol Nowicki-Osuch, based on Nils Eling's code"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here, we perform additional analyses on the similarity between BE and all other tissues.

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(DropletUtils)
library(openxlsx)
library(Rtsne)
library(pheatmap)
library(viridis)
library(ape)
library(batchelor)
source("../../Analysis/Functions/auxiliary.R")

sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Exclude contaminating cells
sce <- sce[,sce$include]

# Exclude duodenum cells
sce <- sce[,colData(sce)$Tissue != "ND"]
```

# Plot circular dendrogram of all clusters

To estimate the similarity between all cell types, we compute an averaged expression profile for each cell-type and perform hierarchical cluster analysis.
Here, we calculate this on the non-batch corrected counts and thereby incorporating the inter-patient variation.
We will first plot the dendogram across all genes, removing lowly expressed genes

```{r}
# Remove non-expressed genes
sce <- sce[Matrix::rowMeans(counts(sce)) > 0,]

# Normalize across all cells
# We provide the cell-types as clusters
sce <- computeSumFactors(sce, clusters=paste(sce$Tissue, sce$cell_type, sep = "_"))
sce <- logNormCounts(sce, log = FALSE)#, return_log = FALSE)

clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(sce))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- Matrix::rowMeans(normcounts(sce)[,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                 sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate spearman correlation between cell types
dend <- hclust(as.dist(sqrt(1 - cor(log2(mat + 1), method = "spearman"))/2), method = "ward.D2")

dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_all_genes.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

Next we will use the same genes as the ones for batch correction.

```{r}
clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")

# Find highly variable genes
# Load sce again to keep internal normalization intact
sce.new <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")
sce.new <- sce.new[,sce.new$include]

# Exclude duodenum cells
sce.new <- sce.new[,colData(sce.new)$Tissue != "ND"]


sce.list <- split.sce(sce.new, unique(sce.new$Sample), colData.name = "Sample")

# Compute the top 1000 shared highly variable genes - use the new code for scran
HVG.genes <- lapply(sce.list, function(n){
  HVG <- modelGeneVar(n)
  HVG
})
  
HVG.df <- do.call("combineVar", HVG.genes)
HVG.df <- HVG.df[order(HVG.df$bio, decreasing = TRUE),]
genes <- rownames(HVG.df)[1:1000]

mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = 1000)
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- Matrix::rowMeans(normcounts(sce)[genes,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                 sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate spearman correlation between cell types
dend <- hclust(as.dist(sqrt(1 - cor(log2(mat + 1), method = "spearman"))/2), method = "ward.D2")

dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_HVG.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

# Multi-batch scaling

We next scale the batches as a way to remove batch effects. 

```{r}
# Read in the data
sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Exclude contaminating cells
sce <- sce[,sce$include]

# Exclude duodenum cells
sce <- sce[,colData(sce)$Tissue != "ND"]



# Split the object into batches
sce.list <- split.sce(sce, unique(sce$Sample), colData.name = "Sample")

# Scale batches
library(batchelor)
func <- paste0("batchelor::multiBatchNorm(", 
                     paste0("sce.list[[", 1:length(sce.list), "]]", collapse=", "), 
                     ")")
sce.scaled <- eval( parse(text=func) )

# Merge these datasets
sce <- do.call("cbind", sce.scaled)

# Remove non-expressed genes
sce <- sce[Matrix::rowMeans(counts(sce)) > 0,]

# Normalize across all cells
# We provide the cell-types as clusters
sce <- computeSumFactors(sce, clusters=paste(sce$Tissue, sce$cell_type, sep = "_"))
sce <- logNormCounts(sce, log = FALSE)
```

We next repeat the analysis as above.
First, we calculate a consensus transcriptome for each cell-type based on all genes.

```{r}
clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(sce))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- Matrix::rowMeans(normcounts(sce)[,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                 sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate spearman correlation between cell types
dend <- hclust(as.dist(sqrt(1 - cor(log2(mat + 1), method = "spearman"))/2), method = "ward.D2")

dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_all_genes.batch_scaled.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

Now with the HVGs.

```{r}
clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")

# Find highly variable genes
# Load sce again to keep internal normalization intact
sce.new <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

sce.new <- sce.new[,sce.new$include]
# Exclude duodenum cells
sce.new <- sce.new[,colData(sce.new)$Tissue != "ND"]


sce.list <- split.sce(sce.new, unique(sce.new$Sample), colData.name = "Sample")

# Compute the top 1000 shared highly variable genes - use the new code for scran
HVG.genes <- lapply(sce.list, function(n){
  HVG <- modelGeneVar(n)
  HVG
})
   
HVG.df <- do.call("combineVar", HVG.genes)
HVG.df <- HVG.df[order(HVG.df$bio, decreasing = TRUE),]
genes <- rownames(HVG.df)[1:1000]

mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = 1000)
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- Matrix::rowMeans(normcounts(sce)[genes,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                 sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate spearman correlation between cell types
dend <- hclust(as.dist(sqrt(1 - cor(log2(mat + 1), method = "spearman"))/2), method = "ward.D2")

dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_HVG.batch_scaled.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

# Different number of HVG

Here, I will use different sets of genes to show that the batch correction is independent from the chosen gene set.

```{r batch-correction}
all_sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

all_sce <- all_sce[,all_sce$include]

# Exclude duodenum cells
all_sce <- all_sce[,colData(all_sce)$Tissue != "ND"]



sce.list <- split.sce(sce = all_sce, groups = unique(all_sce$Sample), colData.name = "Sample")

# Order sce objects for batch correction
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

n <- names(sce.list)
sce.list <- sce.list[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
                       which(grepl("NE_out", n)), which(grepl("NG_out", n)),
                       which(grepl("BE_out", n)), which(grepl("SMG_out", n)))]


# Combine sce objects
all_sce <- do.call("cbind", sce.list)
  
# Batch correction
corrected_1 <- batch.correction(sce.list, number.HVG = 500)
corrected_2 <- batch.correction(sce.list, number.HVG = 2000)

# Compute tsne on corrected counts
set.seed(1234)
# tsne_1 <- Rtsne(t(corrected_1), pca = FALSE)
# tsne_2 <- Rtsne(t(corrected_2), pca = FALSE)

# dendogram analysis
clusters <- paste(colData(all_sce)$Tissue, colData(all_sce)$Tissue_cluster, sep = "_")
mat_1 <- mat_2 <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(corrected_1))
colnames(mat_1) <- colnames(mat_2) <- unique(clusters)

for(i in unique(clusters)){
  mat_1[,i] <- rowMeans(corrected_1[,clusters == i])
  mat_2[,i] <- rowMeans(corrected_2[,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat_1)){
  colnames(mat_1)[i] <- unique(all_sce$cell_type[all_sce$Tissue == sub("_[0-9]*$", "", colnames(mat_1)[i]) &
                                 all_sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat_1)[i])])
  colnames(mat_2)[i] <- unique(all_sce$cell_type[all_sce$Tissue == sub("_[0-9]*$", "", colnames(mat_2)[i]) &
                                 all_sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat_2)[i])])
}

# Calculate euclidean distance on corrected counts
dend_1 <- hclust(dist(t(mat_1), method = "euclidean"), method = "ward.D2")
dend_2 <- hclust(dist(t(mat_2), method = "euclidean"), method = "ward.D2")
dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_500HVG.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend_1), type = "fan", 
     tip.color = metadata(all_sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()

pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_2000HVG.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend_2), type = "fan", 
     tip.color = metadata(all_sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

# Patient correction

Next, I will only correct across the different patients.

```{r}
all_sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")
all_sce <- all_sce[,all_sce$include]


# Exclude duodenum cells
all_sce <- all_sce[,colData(all_sce)$Tissue != "ND"]

sce.list <- split.sce(sce = all_sce, groups = unique(all_sce$Patient), colData.name = "Patient")

# Correctly normalize for each patient
sce.list <- lapply(sce.list, function(n){
  cur_cluster <- quickCluster(n, method = "igraph")#, irlba.args = c("work" = 100))
  n <- computeSumFactors(n, clusters=cur_cluster)
  n <- logNormCounts(n, log = TRUE)
  n
})

# Order list by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]
# "Patient02" "Patient13" "Patient10" "Patient01" "Patient07" "Patient05" "Patient03" "Patient04" "Patient06" "Patient08" "Patient11" "Patient09"


# Combine sce objects
all_sce <- do.call("cbind", sce.list)

# Perform batch correction
corrected <- batch.correction(sce.list)
set.seed(1234)
tsne <- Rtsne(t(corrected), pca = FALSE, perxplexity = 100)



p.tsne.patient.correction <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  clusters = as.factor(colData(all_sce)$Tissue))) + 
  geom_point(aes(tSNE1, tSNE2, colour = clusters)) +
  scale_colour_manual(values = metadata(all_sce)$colour_vector)
p.tsne.patient.correction


# dendogram analysis
clusters <- paste(colData(all_sce)$Tissue, colData(all_sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(corrected))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- rowMeans(corrected[,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(all_sce$cell_type[all_sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                 all_sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate euclidean distance on corrected counts
dend <- hclust(dist(t(mat), method = "euclidean"), method = "ward.D2")
dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_patient_correction.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(all_sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```


# Sample correction wiht duodenum (as in fig 2)

Next, I will correct across all samples with duodenum included

```{r}
sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Exclude contaminating cells
sce <- sce[,sce$include]

# Batch correction
sce.list <- split.sce(sce, unique(sce$Sample), colData.name = "Sample")

# Order sce objects for batch correction
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

n <- names(sce.list)
sce.list <- sce.list[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
                       which(grepl("NE_out", n)), which(grepl("NG_out", n)),
                       which(grepl("ND_out", n)),which(grepl("SMG_out", n)),
                       which(grepl("BE_out", n)))]


corrected <- batch.correction(sce.list)
sce <- do.call("cbind", sce.list)

# Compute new tSNE
set.seed(11111)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 250)
reducedDims(sce)$TSNE <- tsne$Y

# Clustering on corrected data
g <- buildSNNGraph(corrected, k = 10)
clusters <- igraph::cluster_louvain(g)$membership

# Check what clusters overlap with nonepithelial cell types
tmp<-reshape2::dcast(plyr::count(cbind(clusters,sce$tissue_type)), formula = x.clusters ~ x.V2, fill = 0)
tmp<-tmp$x.clusters[tmp$NonEpithelial == apply(as.matrix(tmp[,2:5]), 1, max)]
# View(tmp)
# Exclude immune and stromal clusters
ImmuneTF <- clusters


#Check what cluster overlap with immune and stromal cells in the individual tissues
table(paste(ImmuneTF,colData(sce)$tissue_type, sep = "_"))
ggplot(data.frame(cluster = factor(clusters),
                  Cells = colData(sce)$tissue_type)) +
  geom_bar(aes(fill = Cells, x = cluster), position = "fill", stat = "count")
#10, 13, 19:22, 24
ImmuneTF <- ifelse(ImmuneTF %in% tmp, "nonepi", "epi")


# Visualize clustering with low alpha for immune and stromal cells
ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  clusters = as.factor(clusters),
                  ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tSNE1, tSNE2, colour = clusters, alpha = ImmuneTF), size = 0.5) +
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.05)) 

# Visualize tsne
p.all.cells <- ggplot(data.frame(tsne1 = reducedDims(sce)$TSNE[,1],
                                 tsne2 = reducedDims(sce)$TSNE[,2],
                                 tissue = colData(sce)$Tissue,
                                 ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tsne1, tsne2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.05)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))
p.all.cells
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/All_cells_tsne.pdf", p.all.cells, width = 7, height = 5, useDingbats = FALSE)

#Identify the immune and stromal cells based on the manual annotation
ImmuneTF <- colData(sce)$cell_type
ImmuneTF[!(ImmuneTF %in% c("Immune", "Stromal"))] <- "epi"

# Visualize tsne
ggplot(data.frame(tsne1 = reducedDims(sce)$TSNE[,1],
                                 tsne2 = reducedDims(sce)$TSNE[,2],
                                 tissue = colData(sce)$Tissue,
                                 ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tsne1, tsne2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "Immune" = 0.05, "Stromal" = 0.05)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "grey"))






clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(corrected))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- rowMeans(corrected[,clusters == i])
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                             sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate euclidean distance on corrected counts
dend <- hclust(dist(t(mat), method = "euclidean"), method = "ward.D2")
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])

dev.off()
pdf("/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S12/Tree_corrected_counts_withND.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()

```

# End Matter

To finish get session info:

```{r Endnote, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
sessionInfo()
```
