---
title: "Method validation pseudobulk of scRNA-seq against BE scRNA-seq"
author: "Karol"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

This file presents the data that was created by Figure_S18_Cibersort_prep.Rmd script. It will visualise the similarity between real output and the cell proportion calculated by Cibersort.


```{r setup, include=FALSE}
# export RSTUDIO_PANDOC=/usr/lib/rstudio-server/bin/pando
# Rscript -e "library(rmarkdown); render('Figure_S18_Cibersort_analysis.Rmd')"
knitr::opts_chunk$set(	echo=TRUE,
                       results='asis',
                       fig.align='center')
```

```{r setup, , echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
#Environment setup
library(pheatmap)
library(RColorBrewer)
library(reshape2)
library(ggplot2)
library(knitr)
files.dir<-"~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_figures/S18/"
```


## Relative Results Processing

```{r process_relative, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 11, fig.width= 15}
#read all data from cibersort
ciber.data <- as.matrix(read.delim("~/Dropbox/Postdoc/2019-12-29_BE2020/Results/CIBERSORT/OUT_all/CIBERSORT-Results.txt", row.names = 1))
ciber.data<-ciber.data[order(row.names(ciber.data)),]

removal<-c("P.value","Pearson.Correlation","RMSE", "Absolute.score", "Correlation" )
ciber.data2<-ciber.data[,!(colnames(ciber.data) %in% removal)]

#get information about the cell types
colID<-sapply(colnames(ciber.data2), strsplit, "_")
colID<-data.frame(matrix(unlist(colID), nrow=length(colID), byrow=T))
rownames(colID)<-colnames(ciber.data2)
colnames(colID)<-c("Tissue", "Cell Type")


```

## Relative Results - All

### All cell types all tissues

This image containes a relative score for each cluster fo cell types from each the tissue type. The rows are patients. THe most similar cell types are BE undifferentiated clusters.

```{r, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 11, fig.width= 15}

cells <- read.delim("~/Dropbox/Postdoc/2019-12-29_BE2020/Results/CIBERSORT/Cell_proportions_all.txt", row.names = 1)

cells<-cells[order(rownames(cells)),]

breaksList5=seq(0, ceiling(max((ciber.data2), max((cells)))*40)/40, by = 0.025)


pheatmap((ciber.data2), breaks = breaksList5,color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList5)), main = "(Relative score)", cluster_rows = FALSE, cluster_cols = FALSE)


pheatmap((cells), breaks = breaksList5,color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList5)), main = "(true proportion)", cluster_rows = FALSE, cluster_cols = FALSE)

```


### Correlation between the data

```{r, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 11, fig.width= 15}

tmp<-list()
tmp2<-list()


par(mfrow=c(3,4))
for (i in row.names(cells)) {
  tmp[[i]]<-cor.test(as.numeric((cells[i,colnames(ciber.data2)])), as.numeric((ciber.data2[i,])), method = "pearson", exact = FALSE)
  tmp2[[i]]<-cor(as.numeric((cells[i,colnames(ciber.data2)])), as.numeric((ciber.data2[i,])), method = "pearson")
  plot(as.numeric((cells[i,colnames(ciber.data2)])), as.numeric((ciber.data2[i,])), main = i, ylab = "(Cibersort data)", xlab = "(Real proportion)", xlim = c(0,max(as.numeric((cells[i,])), as.numeric((ciber.data2[i,])))+0.1), ylim = c(0,max(as.numeric((cells[i,])), as.numeric((ciber.data2[i,])))+0.1))
}

par(mfrow=c(1,1))
cor.all<-cor((as.numeric(unlist(cells[,colnames(ciber.data2)]))), as.numeric((ciber.data2)), method = "pearson")

dev.off()
pdf(paste0(files.dir, ".//Figure_S18A.pdf"), useDingbats = FALSE)
plot((as.numeric(unlist(cells[,colnames(ciber.data2)]))), as.numeric((ciber.data2)), main = "All data", ylab = "(Cibersort data)", xlab = "(Real proportion)", xlim = c(0,max((as.numeric(unlist(cells))), as.numeric((ciber.data2)))+0.1), ylim = c(0,max((as.numeric(unlist(cells))), as.numeric((ciber.data2)))+0.1))
dev.off()

kable((unlist(tmp2)))
kable((cor.all), caption = "Global correlation")


```


### Collapsed tissues

This image containes a relative score for each tissue (scores for clusters from each tissue were add)

```{r, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 11, fig.width= 15}

ciberdata4<-as.data.frame(matrix(ncol = 6, nrow = nrow(ciber.data2)))
colnames(ciberdata4)<-c("Sample","BE","NG","NE","SMG", "NonEpi")
ciberdata4$Sample<-rownames(ciber.data2)
ciberdata4$BE<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "BE"]])
ciberdata4$NG<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "NG"]])
ciberdata4$NE<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "NE"]])
ciberdata4$SMG<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "SMG"]])
#ciberdata4$NSCJ<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "NSCJ"]])
ciberdata4$NonEpi<-rowSums(ciber.data2[,colnames(ciber.data2) %in% rownames(colID)[colID$Tissue == "NonEpi"]])

ciberdata4.2<-as.matrix(ciberdata4[,2:6])
rownames(ciberdata4.2)<-ciberdata4$Sample

#collapse cell data
cells4<-as.data.frame(matrix(ncol = 6, nrow = nrow(cells)))
colnames(cells4)<-c("Sample","BE","NG","NE","SMG", "NonEpi")
cells4$Sample<-rownames(cells)
cells4$BE<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "BE"]])
cells4$NG<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "NG"]])
cells4$NE<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "NE"]])
cells4$SMG<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "SMG"]])
# cells4$NSCJ<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "NSCJ"]])
cells4$NonEpi<-rowSums(cells[,colnames(cells) %in% rownames(colID)[colID$Tissue == "NonEpi"]])

cells4.2<-as.matrix(cells4[,2:6])
rownames(cells4.2)<-cells4$Sample

breaksList5=seq(0, ceiling(max((ciberdata4.2), (cells4.2))*40)/40, by = 0.025)





pheatmap((ciberdata4.2), breaks = breaksList5,color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList5)), main = "(Relative score per tissue type)", cluster_rows = FALSE, cluster_cols = FALSE)


pheatmap((cells4.2), breaks = breaksList5,color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList5)), main = "(true proportion per tissue type)", cluster_rows = FALSE, cluster_cols = FALSE)


par(mfrow=c(3,4))
for (i in row.names(cells4.2)) {
  tmp[[i]]<-cor.test(as.numeric((cells4.2[i,])), as.numeric((ciberdata4.2[i,])), method = "pearson", exact = FALSE)
  tmp2[[i]]<-cor(as.numeric((cells4.2[i,])), as.numeric((ciberdata4.2[i,])), method = "pearson")
  
  plot(as.numeric((cells4.2[i,])), as.numeric((ciberdata4.2[i,])), main = i, ylab = "(Ciberdata)", xlab = "(Real proportion)", xlim = c(0,max(as.numeric((cells4.2[i,])), as.numeric((ciberdata4.2[i,])))+0.1), ylim = c(0,max(as.numeric((cells4.2[i,])), as.numeric((ciberdata4.2[i,])))+0.1))
}

cor.all<-cor(as.numeric((cells4.2)), as.numeric((ciberdata4.2)), method = "pearson")
par(mfrow=c(1,1))
plot(as.numeric((cells4.2)), as.numeric((ciberdata4.2)), main = "All data", ylab = "(Ciberdata)", xlab = "(Real proportion)", xlim = c(0,max(as.numeric((cells4.2)), as.numeric((ciberdata4.2)))+0.1), ylim = c(0,max(as.numeric((cells4.2)), as.numeric((ciberdata4.2)))+0.1))


kable((unlist(tmp2)))
kable((cor.all), caption = "Global correlation")

par(mfrow=c(1,1))


ciberdata5<-melt(ciberdata4, ID = "Sample", variable.name = "Tissue", value.name = "Proportion")
ciberdata5$Tissue<-factor(ciberdata5$Tissue, levels = c("NE","SMG","NG","BE","NonEpi")) #NSCJ","ALL"))
```

As above but barchart

```{r, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 11, fig.width= 15}
colour_vector <- vector(length = 8)  #7
names(colour_vector) <- c("NG", "BSCJ", "NE","NSCJ", "BE", "ND","SMG", "NonEpi") 
colour_vector["NG"] <- "#4DBBD5FF"
colour_vector["BSCJ"] <- "#DC0000FF"
colour_vector["NE"] <- "darkred"
colour_vector["NSCJ"] <- "#F39B7FFF"
colour_vector["BE"] <- "#00A087FF"
colour_vector["ND"] <- "#3C5488FF"
colour_vector["SMG"] <- "#B09C85FF"
colour_vector["NonEpi"] <- "grey80"


p<-ggplot(ciberdata5, aes(fill = Tissue, y = Proportion, x = Sample))+geom_bar(stat = "identity")+theme_bw()+theme(axis.text.x = element_text(angle  = 90, hjust = 1)) + ggtitle("Tissue Proportion")+scale_fill_manual(values=colour_vector)
p
ggsave(paste0(files.dir, "./Figure_S18B_1.pdf"), p, width = 15, height = 11)



# dev.off()

cells5<-melt(cells4, ID = "Sample", variable.name = "Tissue", value.name = "Proportion")
cells5$Tissue<-factor(cells5$Tissue, levels = c("NE","SMG","NG","BE","NonEpi"))#,"NSCJ", "ALL"))


p<-ggplot(cells5, aes(fill = Tissue, y = Proportion, x = Sample))+geom_bar(stat = "identity")+theme_bw()+theme(axis.text.x = element_text(angle  = 90, hjust = 1)) + ggtitle("True data per tissue")+scale_fill_manual(values=colour_vector)

p
ggsave(paste0(files.dir, "./Figure_S18B_2.pdf"), p, width = 15, height = 11)



```




## End Matters 

To finish get session info:

```{r Endnote, echo=FALSE}
sessionInfo()
```




