---
title: "IPA_output BE undiff cells vs BE diff"
author: "Karol"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

This is output of the standard analysis of IPA. I am printing z-score of activation for transcriptional regulators with corrected p-value below 0.05. In the IPA, I did standard analysis with the only setting changed to Direct Interactions.

IPA:
Content version: 51963813 (Release Date: 2020-03-11)

```{r setup, include=FALSE}
# export RSTUDIO_PANDOC=/usr/lib/rstudio-server/bin/pando
# Rscript -e "library(rmarkdown); render('FILE_NAME.Rmd')"
knitr::opts_chunk$set(	echo=TRUE,
                       results='asis',
                       fig.align='center')
```

```{r setup, , echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
library(ggplot2)
files.dir<-"/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_4/"
location<-paste0(files.dir, "/Data/IPA/")

```

## Figure 4D top

This is analysis of all the genes transcriptional regulators identified between the undifferentiated and differententiated cells of BE. Only IPA data. 

Content version: 49932394 (Release Date: 2019-11-14)

```{r process_4D, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 14, fig.width= 10}
#Read data
all.data<-read.delim(file=paste0(location,"/IPA_S7_BE_DiffvUndiff_upstream.txt"),  skip = 2)
#Calc log of p-values
all.data$logp<- -log10(all.data$p.value.of.overlap)
# all.data$p.adj<- p.adjust(all.data$p.value.of.overlap, method = "BH")
# all.data$logq<- -log10(all.data$p.adj)
all.data$logq<- -log10(all.data$B.H.corrected.p.value)


all.data<-all.data[rev(order(all.data$Activation.z.score)),]
all.data$no.genes<-sapply(as.character(all.data$Target.Molecules.in.Dataset), function(x) {length(unlist(strsplit(x, split = ",")))})

type = "transcription regulator" 
max=20
data.tmp<-all.data[all.data$Molecule.Type==type,]
data.tmp$Activation.z.score[is.na(data.tmp$Activation.z.score)]<-0

data.tmp<-data.tmp[rev(order(data.tmp$Activation.z.score)),]
#keep only data with corrected p-value below 0.05
data.tmp<-data.tmp[data.tmp$B.H.corrected.p.value<0.05,]
# data.tmp<-data.tmp[abs(data.tmp$Activation.z.score)>1,]


data.tmp$test<-"Other"
data.tmp$test[lapply(data.tmp$Upstream.Regulator, function(x) grep(pattern = "MYC\\b", x, perl = TRUE))>0]<-"MYC"
data.tmp$test[lapply(data.tmp$Upstream.Regulator, function(x) grep(pattern = "HNF4", x))>0]<-"HNF4A"


data.tmp$n<-1:nrow(data.tmp)


p<-ggplot(data.tmp, aes(y= Activation.z.score, x= n, color = test))+geom_point()+scale_color_manual(values = c("#E41A1C" ,"#377EB8", "grey" ))+theme_bw()

p

ggsave(plot = p, file = paste0(files.dir, "./Figure_4DTop.pdf"),  width = 7, height = 6, useDingbats=FALSE)


```

## Figure 4D bottom

Here I am using data from GSEA. 
GSEA software version 4.0.3, C3 TF Database version 7.1



```{r process, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE,fig.height= 14, fig.width= 10}
#Read data
GSEA.dir<-"/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_4/Data/GSEA/BE_undiffvsdiff_C3TFmotifs.GseaPreranked.1585565155707/"

all.data<-read.delim(file=paste0(GSEA.dir,"gsea_report_for_na_pos_1585565155707.xls"))
all.data<-rbind(all.data,read.delim(file=paste0(GSEA.dir,"/gsea_report_for_na_neg_1585565155707.xls")))


all.data$test<-"Other"
all.data$test[lapply(all.data$GS.br..follow.link.to.MSigDB, function(x) grep(pattern = "MYC", x))>0]<-"MYC"
all.data$test[lapply(all.data$GS.br..follow.link.to.MSigDB, function(x) grep(pattern = "HNF4", x))>0]<-"HNF4A"

# all.data$test[is.na(all.data$test)]<-FALSE
all.data<-all.data[rev(order(all.data$NES)),]

all.data$n<-1:nrow(all.data)
all.data<-all.data[rev(order(all.data$test)),]

ggplot(all.data, aes(x= NES, y= FDR.q.val , color = test)) + geom_point()+scale_color_manual(values = c("#E41A1C" ,"#377EB8", "grey" ))+theme_bw()


p<-ggplot(all.data, aes(y= NES, x= n, color = test))+geom_point()+scale_color_manual(values = c("#E41A1C" ,"#377EB8", "grey" ))+theme_bw()


p
ggsave(plot = p, file = paste0(files.dir, "./Figure_4Dbottom.pdf"),  width = 7, height = 6, useDingbats=FALSE)


```


## End Matters

To finish get session info:
```{r Endnote, echo=FALSE}
sessionInfo()
```
