---
title: "Figure 3: Similarity analysis between BE and other tissues"
author: "Karol Nowicki-Osuch, based on Nils Eling's analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script focuses on the similarity analysis between BE and other tissues: NE, NSCJ, BSCJ, NG and SMG.

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(DropletUtils)
library(openxlsx)
library(Rtsne)
library(pheatmap)
library(cowplot)
library(RColorBrewer)
library(edgeR)
library(ape)
library(viridis)
library(umap)
source("../Analysis/Functions/auxiliary.R")
sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Exclude contaminating cells
sce <- sce[,sce$include]

# Exclude duodenum cells
sce <- sce[,colData(sce)$Tissue != "ND"]
```

# Plot tsne

Here, we visualize the batch-corrected counts using tsne.

```{r}
# # Batch correction
# sce.list <- split.sce(sce, unique(sce$Sample), colData.name = "Sample")
# 
# # Order sce objects for batch correction
# sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]
# 
# n <- names(sce.list)
# sce.list <- sce.list[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
#                        which(grepl("NE_out", n)), which(grepl("NG_out", n)),
#                        which(grepl("BE_out", n)), which(grepl("SMG_out", n)))]
# 
# corrected <- batch.correction(sce.list)
# sce <- do.call("cbind", sce.list)
# order the samples by size of the sample 
n <- names(sort(table(sce[["Sample"]]), decreasing = TRUE))
n <- n[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
         which(grepl("NE_out", n)), which(grepl("NG_out", n)),
         which(grepl("BE_out", n)), which(grepl("SMG_out", n)))]
sce<-batchelor::multiBatchNorm(sce, batch = sce[["Sample"]])
# The samples are in NSCJ, NE, NG, SMG order  
corrected <- batch.correction.single(sce, batches = "Sample", m.order = n)



# Compute new tSNE
set.seed(1234)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 250)
umap <-umap(t(corrected))
reducedDims(sce)$TSNE <- tsne$Y
reducedDims(sce)$umap <- umap$layout

# Clustering on corrected data
g <- buildSNNGraph(corrected, k = 10)
clusters <- igraph::cluster_louvain(g)$membership

# Check what clusters overlap with nonepithelial cell types
tmp<-reshape2::dcast(plyr::count(cbind(clusters,sce$tissue_type)), formula = x.clusters ~ x.V2, fill = 0)
tmp<-tmp$x.clusters[tmp$NonEpithelial == apply(as.matrix(tmp[,2:5]), 1, max)]
# View(tmp)
# Exclude immune and stromal clusters
ImmuneTF <- clusters
ImmuneTF <- ifelse(ImmuneTF %in% tmp, "nonepi", "epi")

# Visualize clustering with low alpha for immune and stromal cells
ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  clusters = as.factor(clusters),
                  ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tSNE1, tSNE2, colour = clusters, alpha = ImmuneTF), size = 0.5) +
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.05)) 

# Visualize clustering with low alpha for immune and stromal cells - umap
ggplot(data.frame(UMAP1 = umap$layout[,1],
                  UMAP2 = umap$layout[,2],
                  clusters = as.factor(clusters),
                  ImmuneTF =ImmuneTF)) + 
  geom_point(aes(UMAP1, UMAP2, colour = clusters, alpha = ImmuneTF), size = 0.5) +
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.01)) 

# Visualize tsne
p.all.cells <- ggplot(data.frame(tsne1 = reducedDims(sce)$TSNE[,1],
                                 tsne2 = reducedDims(sce)$TSNE[,2],
                                 tissue = colData(sce)$Tissue,
                                 ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tsne1, tsne2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.05)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )

p.all.cells


# Visualize tsne random order
jittered <- sample(ncol(sce))
p.all.cells_random <- ggplot(data.frame(tsne1 = reducedDims(sce)$TSNE[jittered,1],
                                        tsne2 = reducedDims(sce)$TSNE[jittered,2],
                                        tissue = colData(sce)$Tissue[jittered],
                                        ImmuneTF =ImmuneTF[jittered])) + 
  geom_point(aes(tsne1, tsne2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.05)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )

p.all.cells_random



# Visualize umap
p.all.cells.umap <- ggplot(data.frame(UMAP1 = umap$layout[,1],
                                      UMAP2 = umap$layout[,2],
                                      tissue = colData(sce)$Tissue,
                                      ImmuneTF =ImmuneTF)) + 
  geom_point(aes(UMAP1, UMAP2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(UMAP1, UMAP2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.01)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )

p.all.cells.umap



# Visualize umap in random rder
p.all.cells.umap_random <- ggplot(data.frame(UMAP1 = umap$layout[jittered,1],
                                             UMAP2 = umap$layout[jittered,2],
                                             tissue = colData(sce)$Tissue[jittered],
                                             ImmuneTF =ImmuneTF[jittered])) + 
  geom_point(aes(UMAP1, UMAP2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(UMAP1, UMAP2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "nonepi" = 0.01)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )

p.all.cells.umap_random


#Identify the immune and stromal cells based on the manual annotation
ImmuneTF <- colData(sce)$cell_type
ImmuneTF[!(ImmuneTF %in% c("Immune", "Stromal"))] <- "epi"

# Visualize tsne
ggplot(data.frame(tsne1 = reducedDims(sce)$TSNE[,1],
                  tsne2 = reducedDims(sce)$TSNE[,2],
                  tissue = colData(sce)$Tissue,
                  ImmuneTF =ImmuneTF)) + 
  geom_point(aes(tsne1, tsne2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(tsne1, tsne2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "Immune" = 0.05, "Stromal" = 0.05)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )


# Visualize umap
ggplot(data.frame(UMAP1 = umap$layout[,1],
                  UMAP2 = umap$layout[,2],
                  tissue = colData(sce)$Tissue,
                  ImmuneTF =ImmuneTF)) + 
  geom_point(aes(UMAP1, UMAP2, alpha = ImmuneTF), colour = "black", size = 1) + 
  geom_point(aes(UMAP1, UMAP2, colour = tissue, alpha = ImmuneTF), size = 0.5) + 
  scale_color_manual(values = metadata(sce)$colour_vector) + 
  scale_alpha_manual(values = c("epi" = 1 , "Immune" = 0.01, "Stromal" = 0.01)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.background = element_blank(),
        legend.key = element_blank()
  ) + 
  xlab("t-SNE 1")  + 
  ylab("t-SNE 2") + 
  guides(colour = guide_legend(override.aes = list(size=4), title = "Tissue")
  )

# visualise individual gene expression
gene<-"CHGA"
cur_exp <- ggplot(data.frame(tSNE1 = reducedDims(sce)$TSNE[,1],
                             tSNE2 = reducedDims(sce)$TSNE[,2],
                             clusters = as.factor(colData(sce)$Tissue_cluster),
                             expression = logcounts(sce)[rowData(sce)$Symbol == gene,])) +
  geom_point(aes(tSNE1, tSNE2, colour = expression)) +
  scale_colour_viridis(name = "log2(Expr)") + ggtitle(gene)

print(cur_exp)

gene<-"CHGA"
cur_exp <- ggplot(data.frame(UMAP1 = umap$layout[,1],
                             UMAP2 = umap$layout[,2],
                             clusters = as.factor(colData(sce)$Tissue_cluster),
                             expression = logcounts(sce)[rowData(sce)$Symbol == gene,])) +
  geom_point(aes(UMAP1, UMAP2, colour = expression)) +
  scale_colour_viridis(name = "log2(Expr)") + ggtitle(gene)

print(cur_exp)


gene<-"MUC2"
cur_exp <- ggplot(data.frame(tSNE1 = reducedDims(sce)$TSNE[,1],
                             tSNE2 = reducedDims(sce)$TSNE[,2],
                             clusters = as.factor(colData(sce)$Tissue_cluster),
                             expression = logcounts(sce)[rowData(sce)$Symbol == gene,])) +
  geom_point(aes(tSNE1, tSNE2, colour = expression)) +
  scale_colour_viridis(name = "log2(Expr)") + ggtitle(gene)

print(cur_exp)

gene<-"MUC2"
cur_exp <- ggplot(data.frame(UMAP1 = umap$layout[,1],
                             UMAP2 = umap$layout[,2],
                             clusters = as.factor(colData(sce)$Tissue_cluster),
                             expression = logcounts(sce)[rowData(sce)$Symbol == gene,])) +
  geom_point(aes(UMAP1, UMAP2, colour = expression)) +
  scale_colour_viridis(name = "log2(Expr)") + ggtitle(gene)

print(cur_exp)

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/All_cells_noND.pdf", p.all.cells, width = 7, height = 5, useDingbats = FALSE)

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/All_cells_noND_umap.pdf", p.all.cells.umap, width = 7, height = 5, useDingbats = FALSE)

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/All_cells_noND_random.pdf", p.all.cells_random, width = 7, height = 5, useDingbats = FALSE)

ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/All_cells_noND_umap_random.pdf", p.all.cells.umap_random, width = 7, height = 5, useDingbats = FALSE)
```

# Plot circular dendrogram of all clusters

To estimate the similarity between all cell types, we compute an averaged expression profile for each cell-type and perform hierarchical cluster analysis.

```{r}
# We will perform this analysis on the corrected and uncorrected counts (uncorrected in supplements)
clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(corrected))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- rowMeans(corrected[,clusters == i])
}

# # Alternative calculation of average consensus
# for(i in unique(clusters)){
#   tmp<-apply(corrected[,clusters == i], 1, function (x) {aggregate(x, by  = list(colData(sce)$Sample[clusters ==i]), mean)})
#   mat[,i] <- unlist(lapply(tmp, function(x){mean(x$x)}))
# }

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                             sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate euclidean distance on corrected counts
dend <- hclust(dist(t(mat), method = "euclidean"), method = "ward.D2")
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])

dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/Tree_corrected_counts_noND.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```


## Plot circular dendrogram of all clusters using highly variable genes

To estimate the similarity between all cell types, we compute an averaged expression profile for each cell-type and perform hierarchical cluster analysis.

```{r, eval = FALSE}
# We will perform this analysis on the corrected and uncorrected counts (uncorrected in supplements)
HVG.genes<-modelGeneVar(sce, block = sce[["Sample"]])
genes<- getTopHVGs(HVG.genes, n = 1000)


clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = 1000)
colnames(mat) <- unique(clusters)

# for(i in unique(clusters)){
#   mat[,i] <- rowMeans(corrected[,clusters == i])
# }

# Alternative calculation of average consensus
for(i in unique(clusters)){
  tmp<-apply(logcounts(sce)[genes,clusters == i], 1, function (x) {aggregate(x, by  = list(colData(sce)$Sample[clusters ==i]), mean)})
  mat[,i] <- unlist(lapply(tmp, function(x){mean(x$x)}))
}

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                             sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate euclidean distance on corrected counts
# dend <- hclust(dist(t(mat), method = "euclidean"), method = "ward.D2")
dend <- hclust(as.dist(sqrt(1 - cor(mat, method = "spearman"))/2), method = "ward.D2")

plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])


pheatmap(mat, scale = "row", clustering_distance_cols = "correlation", clustering_method = "ward.D2")

# dev.off()
# pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/Tree_corrected_counts_noND.pdf", width = 10, height = 10, useDingbats = FALSE)
# plot(as.phylo(dend), type = "fan", 
#      tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
# dev.off()
```

# Plot circular dendrogram of all clusters - alternative names

To estimate the similarity between all cell types, we compute an averaged expression profile for each cell-type and perform hierarchical cluster analysis.

```{r}
# We will perform this analysis on the corrected and uncorrected counts (uncorrected in supplements)
clusters <- paste(colData(sce)$Tissue, colData(sce)$Tissue_cluster, sep = "_")
mat <- matrix(data = NA, ncol = length(unique(clusters)), nrow = nrow(corrected))
colnames(mat) <- unique(clusters)

for(i in unique(clusters)){
  mat[,i] <- rowMeans(corrected[,clusters == i])
}

# # Alternative calculation of average consensus
# for(i in unique(clusters)){
#   tmp<-apply(corrected[,clusters == i], 1, function (x) {aggregate(x, by  = list(colData(sce)$Sample[clusters ==i]), mean)})
#   mat[,i] <- unlist(lapply(tmp, function(x){mean(x$x)}))
# }

# Rename the matrix to contain the actual cell-type labels
for(i in 1:ncol(mat)){
  colnames(mat)[i] <- unique(sce$cell_type_secondary[sce$Tissue == sub("_[0-9]*$", "", colnames(mat)[i]) &
                                                       sce$Tissue_cluster == sub("^[A-Z2]*_", "", colnames(mat)[i])])
}

# Calculate euclidean distance on corrected counts
dend <- hclust(dist(t(mat), method = "euclidean"), method = "ward.D2")
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])


dev.off()
pdf("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Figure_3/Tree_corrected_counts_noND_alt_names.pdf", width = 10, height = 10, useDingbats = FALSE)
plot(as.phylo(dend), type = "fan", 
     tip.color = metadata(sce)$colour_vector[sub("_[0-9]*$", "", unique(clusters))])
dev.off()
```

# End Matter

To finish get session info:

```{r Endnote, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
sessionInfo()
```
