---
title: "Rebuttal Round two"
author: "Karol Nowicki-Osuch"
date: "`r Sys.Date()`"
output: 
html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here I am visualising different aspects of the analysis. First let's take a look at the trajectories between the samples.

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(batchelor)
# library(DropletUtils)
# library(openxlsx)
library(Rtsne)
# library(pheatmap)
# library(cowplot)
# library(RColorBrewer)
# library(edgeR)
# library(ape)
# library(viridis)
# library(umap)
# library(reshape2)
library(TSCAN)
library(destiny)
# library(anndata)
source("./Analysis/Functions/auxiliary.R")


```

# Perform pseudotime analysis

Here I will use TSCAN algorithm to reconstruct the trajectory following this manual: https://bioconductor.org/books/release/OSCA/trajectory-analysis.html


```{r}
# Read in all data
sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")

# Exclude contaminating cells
sce <- sce[,sce$include]

# Exclude duodenum cells
sce <- sce[,colData(sce)$Tissue != "ND"]

# order the samples by size of the sample 
n <- names(sort(table(sce[["Sample"]]), decreasing = TRUE))
n <- n[c(which(grepl("NSCJ_out", n)), which(grepl("BSCJ_out", n)), 
         which(grepl("NE_out", n)), which(grepl("NG_out", n)),
         which(grepl("BE_out", n)), which(grepl("SMG_out", n)))]
sce<-batchelor::multiBatchNorm(sce, batch = sce[["Sample"]])
# The samples are in NSCJ, NE, NG, SMG order  
corrected <- batch.correction.single(sce, batches = "Sample", m.order = n)


# Perform PCA
HVG.genes<-modelGeneVar(sce, block = sce[["Sample"]])
top.sce <- getTopHVGs(HVG.genes, n=1000)

set.seed(1234)
PCA <- prcomp(t(corrected))

# save a slot
reducedDims(sce)$PCA <- PCA$x

plot(summary(PCA)$importance[2,])

# Get new t-SNE on this machine
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 250)

reducedDims(sce)$TSNE <- tsne$Y

plotReducedDim(sce, dimred="TSNE", colour_by="Tissue")


#Aggregate data across tissue clusters
by.cluster<-aggregateAcrossCells(sce[,sce$tissue_type != "NonEpithelial"], ids = paste(sce$Tissue[sce$tissue_type != "NonEpithelial" ], sce$Tissue_cluster[sce$tissue_type != "NonEpithelial"] , sep = "_"))

# Get the centers of each cluster in the TSNE space
centroids <- reducedDim(by.cluster, "PCA")

mst <- TSCAN::createClusterMST(centroids, clusters=NULL)

line.data <- reportEdges(by.cluster, mst=mst, clusters=NULL, use.dimred="TSNE")

plotTSNE(sce[,sce$tissue_type != "NonEpithelial"], colour_by="Tissue") + 
  geom_line(data=line.data, mapping=aes(x=dim1, y=dim2, group=edge))

```

# FLE projection

In additional to PCA and T-SNE projection, I also looked at the force layout embedding. This is actually quite good at conveing the timing of gene expression in temporal samples. Although we do not have temporal data, we stil have developmental data (within each tissue) and these plots are nice visually. I will use the guide from https://broadinstitute.github.io/wot/tutorial/ to plot. it is based on this paper: https://doi.org/10.1016/j.cell.2019.01.006

## get matrix of HVG

```{r}
mat<-counts(sce)
colnames(mat) <- paste(sce$Patient, sce$Tissue, sce$Barcode, sep = "_")
mat.hvg <- mat[top.sce,]
ids<-cbind(paste(sce$Patient, sce$Tissue, sce$Barcode, sep = "_"), sce$Patient)
colnames(ids)<-c("id", "patient")

# write.csv(as.matrix(mat.hvg), file = "~/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/hvg_all.csv")

sceasy:::convertFormat(sce[top.sce,], from="sce", to="anndata",
                         outFile="/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/hvg_all.h5ad")

write.table(ids, file = "~/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/ids_all.txt", sep = "\t", quote = FALSE, row.names = FALSE)

```

## Run python scripts to get embedding

```{python, eval = FALSE}
import numpy as np
import pandas as pd
from matplotlib import pyplot as plt
import pegasus as pg
import anndata as an
import wot

# Path to input files
# FULL_DS_PATH = 'Data/all.csv'
# VAR_DS_PATH = '/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/hvg_all.csv'
VAR_DS_PATH = '/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/hvg_all.h5ad'
IDS_PATH = '~/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/ids_all.txt'

ids_df = pd.read_csv(IDS_PATH, index_col='id', sep='\t')

#adata = wot.io.read_dataset(FULL_DS_PATH, obs=[ids_df])
adata_var = wot.io.read_dataset(VAR_DS_PATH, obs=[ids_df])

pg.pca(adata_var, features=None)
pg.neighbors(adata_var)
pg.diffmap(adata_var)

pg.fle(adata_var, n_jobs=24)
coords = adata_var.obsm['X_fle']
pd.DataFrame(coords).to_csv("~/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/coords_all.csv")





```

## Load Embedding


```{r}
fle<-read.csv("~/Dropbox/Postdoc/2019-12-29_BE2020/Trajectory/FLE/Data/coords_all.csv")
reducedDims(sce)$fle <- fle[,2:3]

plotReducedDim(sce[,sce$tissue_type != "NonEpithelial"], dimred="fle", colour_by="Tissue")
```

# Diffusion map

From Destiny package

```{r}
# Run destiny
dm <- DiffusionMap(as.matrix(logcounts(sce)[top.sce,]))

plot(dm, 1:2)


# dm@eigenvectors[,1:2]
reducedDims(sce)$dm <- dm@eigenvectors[,1:2]


```

# run TSCAN

```{r}
# Identify non epithelial clusters
g <- buildSNNGraph(corrected, k = 5)
clusters <- igraph::cluster_louvain(g)$membership
good.clusters <- c(1:max(clusters))[table(factor(clusters[sce$tissue_type != "NonEpithelial"], levels = 1:max(clusters) ))/table(clusters)>0.95]


# Agregate data across cluster (actually detailed description of cell types)
by.cluster<-aggregateAcrossCells(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters], ids = sce$cell_type_secondary[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])

# Run TSCAN on PCA data with outgroup option on and wiht cell types as clusters and with mnn mode on. Only focus on epithelial cells
pseudo.mnn <- TSCAN::quickPseudotime(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters], use.dimred="PCA", outgroup=TRUE, clusters =  sce$cell_type_secondary[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters], with.mnn = TRUE)

mnn.pseudo <- rowMeans(pseudo.mnn$ordering, na.rm=TRUE)

set.seed(10101)
line.data <- reportEdges(by.cluster, mst=pseudo.mnn$mst, clusters=NULL, use.dimred="dm")

DM.plot<-ggplot(data.frame(dm1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$dm[,1],
                  dm2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$dm[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(dm1, dm2, colour = Cell), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=DC1, y=DC2, group=edge)) + theme_minimal()

DM.plot
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/DM.plot.pdf", DM.plot, width = 15, height = 5, useDingbats = FALSE)

DM.plot.tissue<-ggplot(data.frame(dm1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$dm[,1],
                  dm2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$dm[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(dm1, dm2, colour = Tissue), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=DC1, y=DC2, group=edge)) + theme_minimal()

DM.plot.tissue
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/DM.plot.tissue.pdf", DM.plot.tissue, width = 10, height = 5, useDingbats = FALSE)


line.data <- reportEdges(by.cluster, mst=pseudo.mnn$mst, clusters=NULL, use.dimred="TSNE")

TSNE.plot<-ggplot(data.frame(tsne1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$TSNE[,1],
                  tsne2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$TSNE[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(tsne1, tsne2, colour = Cell), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=dim1, y=dim2, group=edge)) + theme_minimal()

TSNE.plot
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/TSNE.plot.pdf", TSNE.plot, width = 15, height = 5, useDingbats = FALSE)

TSNE.plot.tissue<-ggplot(data.frame(tsne1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$TSNE[,1],
                  tsne2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$TSNE[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(tsne1, tsne2, colour = Tissue), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=dim1, y=dim2, group=edge)) + theme_minimal()

TSNE.plot.tissue
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/TSNE.plot.tissue.pdf", TSNE.plot.tissue, width = 10, height = 5, useDingbats = FALSE)

line.data <- reportEdges(by.cluster, mst=pseudo.mnn$mst, clusters=NULL, use.dimred="fle")

FLE.plot<-ggplot(data.frame(fle1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$fle[,1],
                  fle2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$fle[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(fle1, fle2, colour = Cell), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=X0, y=X1, group=edge)) + theme_minimal()

FLE.plot
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/FLE.plot.pdf", FLE.plot, width = 15, height = 5, useDingbats = FALSE)

FLE.plot.tissue<-ggplot(data.frame(fle1 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$fle[,1],
                  fle2 = reducedDims(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters])$fle[,2],
                  Tissue = as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$Tissue),
                  Cluster = as.factor(clusters[!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]),
                  Cell =  as.factor(sce[,!(sce$cell_type %in% c("Immune", "Stromal")) & clusters %in% good.clusters]$cell_type_secondary)
)) + 
  geom_point(aes(fle1, fle2, colour = Tissue), size = 0.5) + 
  geom_line(data=line.data, mapping=aes(x=X0, y=X1, group=edge)) + theme_minimal()

FLE.plot.tissue
ggsave("~/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Rebuttal/FLE.plot.tissue.pdf", FLE.plot.tissue, width = 10, height = 5, useDingbats = FALSE)



```

# End Matter

To finish get session info:

```{r Endnote, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
sessionInfo()
```


