---
title: "Figure 1: overview on experimental approach and gland cells"
author: "Karol Nowicki-Osuch based on Nils Eling's code"
date: "`r Sys.Date()`"
output: 
html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script visualizes the full dataset as well as highlighting the sub-populations in the gland cells.

# Read in the data

```{r libraries, message=FALSE}
library(scran)
library(scater)
library(DropletUtils)
library(openxlsx)
library(Rtsne)
library(pheatmap)
library(viridis)
library(umap)

source("./Analysis/Functions/auxiliary.R")

# Read in the normalized and batch-corrected reads
sce <- readRDS("~/Dropbox/Postdoc/2019-12-29_BE2020/All_corrected_sce_filtered.rds")
```

# Visualize gland cells

To visualize the gland cells, we perform batch correction across all patients and colour cells based on the clustering performed in the Tissue_correction script.

```{r}
cur_sce <- sce[,sce$Tissue == "SMG" & sce$include]

# Perform batch scaling
cur_sce<-batchelor::multiBatchNorm(cur_sce, batch = cur_sce[["Sample"]])
batch.order<-order(table(cur_sce[["Sample"]]), decreasing = TRUE)
corrected <- batch.correction.single(cur_sce, batches = "Sample", m.order = batch.order)



# Remove non-expressed genes
# cur_sce <- cur_sce[Matrix::rowSums(counts(cur_sce)) > 0,]


genes <- c("KRT7", "KRT8", "KRT5", "KRT23",
           "MUC5B", "MUC5AC", "MUC2",
           "WFDC2", "KRT19", "SOX9", "TP63")

# Remove somatic cells
cur_corrected <- corrected[,cur_sce$cell_type %in% 
                             c("Mucous", "Oncocytes", "Duct_Intercalating", "Myo-epithelial")]
sel_sce <- cur_sce[,cur_sce$cell_type %in% 
                     c("Mucous", "Oncocytes", "Duct_Intercalating", "Myo-epithelial")]
levels(sel_sce$cell_type)
sel_sce <- logNormCounts(sel_sce)
for.heatmap <- logcounts(sel_sce)[match(genes, rowData(sel_sce)$Symbol),]
colnames(for.heatmap) <- paste(sel_sce$Barcode, sel_sce$Patient, sep = "_")

# Calculate euclidean distance using the factors after batch correction
euc.dist <- dist(t(cur_corrected), method = "euclidean")

# Colouring for cell type
colour_vector <-  vector(length = length(unique(sel_sce$cell_type)))
names(colour_vector) <- unique(sel_sce$cell_type)
# colour_vector["Immune"] <- "grey40"
# colour_vector["Unknown.Doublets"] <- "white"
colour_vector["Mucous"] <- "saddlebrown"
colour_vector["Oncocytes"] <- "burlywood3"
# colour_vector["Stromal"] <- "grey60"
# colour_vector["Nonepithelial"] <- "grey80"
colour_vector["Duct_Intercalating"] <- "burlywood4"
# colour_vector["Squamous_Esophagus"] <- "black"
colour_vector["Myo-epithelial"] <- "brown"
# colour_vector["Dendritic"] <- "grey20"

# Colouring for patient
patient_vector <- c("dark red", "dark blue", "dark green")
names(patient_vector) <- c("Patient10", "Patient11", "Patient13")
pheatmap(for.heatmap, cluster_rows = FALSE, color = viridis(100), 
         labels_row = genes, show_colnames = FALSE, clustering_distance_cols = euc.dist,
         annotation_col = data.frame(row.names = colnames(for.heatmap),
                                     cell_type = sel_sce$cell_type,
                                     patient = sel_sce$Patient), 
         annotation_colors = list(cell_type = colour_vector,
                                  patient = patient_vector), 
         clustering_method = "ward.D2")
dev.off()

pdf(file = "/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_Note/1_SMG_heatmap_lognorm.pdf", height = 7, width = 8, useDingbats = FALSE)
pheatmap(for.heatmap, cluster_rows = FALSE, color = viridis(100), 
         labels_row = genes, show_colnames = FALSE, clustering_distance_cols = euc.dist,
         annotation_col = data.frame(row.names = colnames(for.heatmap),
                                     cell_type = sel_sce$cell_type,
                                     patient = sel_sce$Patient), 
         annotation_colors = list(cell_type = colour_vector,
                                  patient = patient_vector), 
         clustering_method = "ward.D2")
dev.off()
```


# Barret's oesophagus

We will visualize additional marker genes of Barret's oesophagus.

## Dimensionality reduction

```{r}
cur_sce <- sce[,sce$Tissue == "BE" & sce$include]

# Only genes that are expressed
cur_sce <- cur_sce[Matrix::rowSums(logcounts(cur_sce)) > 0,]

# Normalization across all cells
cur_sce <- computeSumFactors(cur_sce, clusters=paste(cur_sce$cell_type_secondary))
cur_sce <- logNormCounts(cur_sce, log = TRUE)


# Split the samples to perform batch correction
sce.list <- split.sce(cur_sce, unique(cur_sce$Sample), colData.name = "Sample")
# Order sce object by size
sce.list <- sce.list[order(unlist(lapply(sce.list, ncol)), decreasing = TRUE)]

cur_sce <- do.call("cbind", sce.list)
corrected <- batch.correction(sce.list)

# Compute tSNE
set.seed(12345)
tsne <- Rtsne(t(corrected), pca = FALSE, perplexity = 100)
umap <- umap(t(corrected))

# Create colour vector
colour_vector <- vector(length = length(unique(cur_sce$cell_type_secondary)))
names(colour_vector) <- unique(cur_sce$cell_type_secondary)
colour_vector["Immune_T-cells"] <- "grey20"
colour_vector["Immune_B-cells"] <- "grey30"
colour_vector["Immune_Macrophages"] <- "grey40"
colour_vector["Stromal_CALD1_cells"] <- "grey60"
colour_vector["Stromal_GNG11_cells"] <- "grey70"

colour_vector["Columnar_Undifferentiated"] <- "white"
colour_vector["Columnar_Undifferentiated_Dividing"] <- colorRampPalette(c("white", "#00A087FF"))(10)[2]
colour_vector["Columnar_Intermediate"] <- colorRampPalette(c("white", "#00A087FF"))(10)[4]
colour_vector["Columnar_differentiated"] <- colorRampPalette(c("white", "#00A087FF"))(10)[6]
colour_vector["Endocrine_NEUROG3"] <- colorRampPalette(c("white", "dark green"))(10)[5]
colour_vector["Goblet"] <- colorRampPalette(c("white", "dark green"))(10)[10]
colour_vector["Endocrine"] <- colorRampPalette(c("white", "dark green"))(10)[5]

tsne.BE <- ggplot(data.frame(tSNE1 = tsne$Y[,1],
                  tSNE2 = tsne$Y[,2],
                  cell_type = cur_sce$cell_type_secondary)) + 
  geom_point(aes(tSNE1, tSNE2), colour = "black", size = 2) +
  geom_point(aes(tSNE1, tSNE2, colour = cell_type)) +
  theme_minimal()+
  scale_color_manual(values = colour_vector)
tsne.BE

gene1<-"MUC5AC"
MUC5ACcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size = 1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
MUC5ACcells



gene1<-"OLFM4"
OLFM4cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
OLFM4cells



gene1<-"FABP1"
FABP1cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
FABP1cells

gene1<-"KLF4"
KLF4cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
KLF4cells

gene1<-"NEUROG3"
NEUROG3cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
NEUROG3cells

gene1<-"CHGA"
CHGAcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
CHGAcells


gene1<-"CDX1"
CDX1cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
CDX1cells


gene1<-"CDX2"
CDX2cells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
CDX2cells


gene1<-"HNF4A"
HNF4Acells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
HNF4Acells

gene1<-"MYC"
MYCcells<-ggplot(data.frame(tSNE1 = tsne$Y[,1],
                              tSNE2 = tsne$Y[,2],
                              gene = logcounts(cur_sce)[rowData(cur_sce)$Symbol == gene1,]))  +
  geom_point(aes(tSNE1, tSNE2, colour = gene), size =1) + 
  scale_colour_viridis(option = "A", name = "log2(Expr)", ) +
  guides(alpha=FALSE) + ggtitle(gene1) + 
  theme_minimal()
MYCcells

# lay <- rbind(c(1,1,1,2),
#              c(1,1,1,3),
#              c(4,5,6,7)
# )
lay <- rbind(c(1,1,1),
             c(1,1,1),
             c(1,1,1),
             c(2,3,4),
             c(5,6,7)
)

all.figures <- grid.arrange(tsne.BE, MUC5ACcells, FABP1cells, OLFM4cells, NEUROG3cells, CHGAcells,KLF4cells, layout_matrix=lay)
ggsave("/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_Note/2_BE_cells_tsne.pdf", all.figures, width = 10, height = 14, useDingbats = FALSE)


lay <- rbind(c(1,1,1),
             c(1,1,1),
             c(1,1,1),
             c(2,3,4),
             c(5,6,7),
             c(8,9,10)
)

all.figures <- grid.arrange(tsne.BE, MUC5ACcells, FABP1cells, OLFM4cells, NEUROG3cells, CHGAcells,KLF4cells, CDX2cells, HNF4Acells, MYCcells, layout_matrix=lay)
ggsave("/home/karolno/Dropbox/Postdoc/2019-12-29_BE2020/Figures/Supplementary_Note/2_BE_cells_tsne_2.pdf", all.figures, width = 14, height = 20, useDingbats = FALSE)


```


# End Matter

To finish get session info:

```{r Endnote, include=TRUE, echo=TRUE, warning=FALSE, eval=TRUE, message=FALSE, tidy=TRUE}
sessionInfo()
```


